---
title: "Crop Value Viz"
---

In this RMD file I will run an initial EDA to compare the crops.

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(ggcorrplot)
library(corrplot)
```

```{r}
setwd('C:/Users/donse/OneDrive/Documents/Fall2021/Case_Studies/Case-Studies-Case-2-Team-1/data')
# Load data
Crop_Details_Standardized <- read_csv("Crop_Details_Standardized.csv")
```

```{r}
crop_cost <- Crop_Details_Standardized %>%
  select(Crop, `Cost of cultivation
(in INR) (per 6 by 6 m area)`)
crop_cost$type <- "cost"
colnames(crop_cost) <- c("crop", "value", "type")

crop_profit <- Crop_Details_Standardized %>%
  select(Crop, `Profit
(in INR) (per 6 by 6 m area)`)
crop_profit$type <- "profit"
colnames(crop_profit) <- c("crop", "value", "type")

value_plot_df <- rbind(crop_cost, crop_profit)

```

```{r}
ggplot(value_plot_df, aes(x = reorder(crop,-value), y = value, fill = type)) +
  theme_bw()+
  geom_bar(stat = "identity", position = 'dodge') + 
  scale_y_continuous(name = "Rupees", breaks = seq(0,5000,1000), labels = function(x) paste0(x, "₹"), limits = c(0, 5700)) +
  scale_fill_manual(breaks = c("cost", "profit"), 
                    values=c("black", "darkgreen")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   labs(
      title = "Comparing the Costs and The Profits Of Growing Each Crop",
      subtitle = "in a 6m by 6m plot",
      x = "Crop",
      y = "Number of Rupees"
   )
```

Interesting. This plot show us the most expensive and the most profitable crops. Now I would like to explore the gap between cost and profit for each crop. Some crops might make more money but cost more to 
cultivate.

```{r}
margin_df <- Crop_Details_Standardized %>%
  mutate(margin = `Profit
(in INR) (per 6 by 6 m area)` - `Cost of cultivation
(in INR) (per 6 by 6 m area)`) %>%
  select(Crop, margin)
```
 
```{r}
ggplot(margin_df, aes(x = reorder(Crop,-margin), y = margin)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_y_continuous(name = "Margin Between Cost and Profit", breaks = seq(0,5000,1000), labels = function(x) paste0(x, "₹"),       limits = c(0, 5000)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
    labs(
       title = "Comparing the gaps between Cost and Profit Each Crop",
       subtitle = "in a 6m by 6m plot",
       x = "crop") 
  
```

Now one last plot, we would like to look at the proportion of the profit that each crops cost to plant is.

```{r}
proportion_df <- Crop_Details_Standardized %>%
  mutate(proportion = `Profit
(in INR) (per 6 by 6 m area)` / `Cost of cultivation
(in INR) (per 6 by 6 m area)`) %>%
  select(Crop, proportion)
```

```{r}
ggplot(proportion_df, aes(x = reorder(Crop,-proportion), y = proportion)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_y_continuous(name = "Times greater", breaks = seq(0,10,2), labels = function(x) paste0(x, "x"),       limits = c(0, 11)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
    labs(
       title = "How Many Times Greater is the Profit than the Cost To Plant",
       subtitle = "for each crop in a 6m by 6m plot",
       x = "crop") 
  
```

