---
title: "Crop VIZ"
---

In this RMD file I will run an initial EDA to compare the crops.

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(ggcorrplot)
library(corrplot)
```

```{r}
setwd('C:/Users/donse/OneDrive/Documents/Fall2021/Case_Studies/Case-Studies-Case-2-Team-1/data')
# Load data
crop_df <- read_csv("Crop_recommendation.csv")
standardized_crop_df <- read_csv("Crop_recommendation_standardized.csv")
```

```{r}
crop_df %>%
  group_by(label) %>%
  summarise(weather_var = var(rainfall) + var(temperature) + var(humidity),
            soil_var = var(N) + var(P) + var(K) + var(ph)) %>%
  arrange(-weather_var)
```

```{r}
crop_comp_df <- standardized_crop_df %>%
  group_by(label) %>%
  summarise(weather_var = var(rainfall) + var(temperature) + var(humidity),
            soil_var = var(N) + var(P) + var(K) + var(ph)) %>%
  arrange(-soil_var)
```

```{r}

weather_df <- crop_comp_df %>%
  select(label, weather_var)
colnames(weather_df) <- c("label", "var")
weather_df$type <- "weather"

soil_df <- crop_comp_df %>%
  select(label, soil_var)
colnames(soil_df) <- c("label", "var")
soil_df$type <- "soil"

plot_df <- rbind(weather_df,soil_df) 

```


```{r}
ggplot(plot_df, aes(x = reorder(label,-var), y = var, fill = type)) +
  theme_bw()+
  geom_bar(stat = "identity", position = 'dodge') + 
   # scale_y_continuous(name = "Percentage of Total Population", breaks = seq(0,100,20), labels = function(x)                                    paste0(x, "%"), limits = c(0, 100)) +
  scale_fill_manual(breaks = c("weather", "soil"), 
                    values=c("turquoise1", "saddlebrown")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   labs(
      title = "Comparing the Variance of Soil Variables and Weather Variables",
      subtitle = "for each different crop",
      x = "crop",
      y = "Variance"
   )
```


```{r}
fullweather_df <- standardized_crop_df %>%
  group_by(label) %>%
  summarise(rain_var = var(rainfall), 
            temp_var = var(temperature),
            hum_var = var(humidity))

rain_df <- fullweather_df %>%
  select(label, rain_var)
colnames(rain_df) <- c("label", "var")
rain_df$type <- "rain"

temp_df <- fullweather_df %>%
  select(label, temp_var)
colnames(temp_df) <- c("label", "var")
temp_df$type <- "temperature"

hum_df <- fullweather_df %>%
  select(label, hum_var)
colnames(hum_df) <- c("label", "var")
hum_df$type <- "humidity"

weather_plot_df <- rbind(rain_df, temp_df, hum_df)

```


```{r}
ggplot(weather_plot_df, aes(x = reorder(label,-var), y = var, fill = type)) +
  theme_bw()+
  geom_bar(stat = "identity", position = 'dodge') + 
   # scale_y_continuous(name = "Percentage of Total Population", breaks = seq(0,100,20), labels = function(x)                                    paste0(x, "%"), limits = c(0, 100)) +
  scale_fill_manual(breaks = c("rain", "temperature", "humidity"), 
                    values=c("turquoise1", "saddlebrown", "darkorange1")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   labs(
      title = "Comparing the Variance of the Weather Variables",
      subtitle = "for each different crop",
      x = "crop",
      y = "Variance"
   )
```


```{r}
fullsoil_df <- standardized_crop_df %>%
  group_by(label) %>%
  summarise(N_var = var(N),
            P_var = var(P), 
            K_var = var(K), 
            ph_var = var(ph)) 

N_df <- fullsoil_df %>%
  select(label, N_var)
colnames(N_df) <- c("label", "var")
N_df$type <- "N"

P_df <- fullsoil_df %>%
  select(label, P_var)
colnames(P_df) <- c("label", "var")
P_df$type <- "P"

K_df <- fullsoil_df %>%
  select(label, K_var)
colnames(K_df) <- c("label", "var")
K_df$type <- "K"

ph_df <- fullsoil_df %>%
  select(label, ph_var)
colnames(ph_df) <- c("label", "var")
ph_df$type <- "ph"

soil_plot_df <- rbind(N_df, P_df, K_df, ph_df)
 
```

```{r}
ggplot(soil_plot_df, aes(x = reorder(label,-var), y = var, fill = type)) +
  theme_bw()+
  geom_bar(stat = "identity", position = 'dodge') + 
   # scale_y_continuous(name = "Percentage of Total Population", breaks = seq(0,100,20), labels = function(x)                                    paste0(x, "%"), limits = c(0, 100)) +
  scale_fill_manual(breaks = c("N", "P", "K", "ph"), 
                    values=c("turquoise1", "saddlebrown", "darkorange1", "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   labs(
      title = "Comparing the Variance of the Soil Variables",
      subtitle = "for each different crop",
      x = "crop",
      y = "Variance"
   )


```

Wow it seems like ph varies wildly for many crops but the rest of the soil variables do not let's remake the plot above one time without ph and see

```{r}
fullsoil_df <- standardized_crop_df %>%
  group_by(label) %>%
  summarise(N_var = var(N),
            P_var = var(P), 
            K_var = var(K), 
            ph_var = var(ph)) 

N_df <- fullsoil_df %>%
  select(label, N_var)
colnames(N_df) <- c("label", "var")
N_df$type <- "N"

P_df <- fullsoil_df %>%
  select(label, P_var)
colnames(P_df) <- c("label", "var")
P_df$type <- "P"

K_df <- fullsoil_df %>%
  select(label, K_var)
colnames(K_df) <- c("label", "var")
K_df$type <- "K"

soil_plot_df <- rbind(N_df, P_df, K_df)
 
```


```{r}
ggplot(soil_plot_df, aes(x = reorder(label,-var), y = var, fill = type)) +
  theme_bw()+
  geom_bar(stat = "identity", position = 'dodge') + 
   # scale_y_continuous(name = "Percentage of Total Population", breaks = seq(0,100,20), labels = function(x)                                    paste0(x, "%"), limits = c(0, 100)) +
  scale_fill_manual(breaks = c("N", "P", "K"), 
                    values=c("turquoise1", "saddlebrown", "darkorange1")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   labs(
      title = "Comparing the Variance of the Soil Variables",
      subtitle = "for each different crop",
      x = "crop",
      y = "Variance"
   )


```

All of these variances are small but K is the smallest. P is slightly bigger and even though N is the largest it is still relatively small. For my last plot I would like to make a corr plot to show the correlations between all of the variables. 

```{r}
cor_plot_df <- standardized_crop_df[,1:7]
cor_plot_df <- cor_plot_df %>%
  na.omit()

data(cor_plot_df)
cor_plot_df <- cor(cor_plot_df)

corrplot(cor_plot_df, type = "lower")
```

```{r}
ggcorrplot(cor_plot_df, hc.order = TRUE,
           hc.method = method,
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of Weather Variables and Soil Variables for different Crops",
           ggtheme=theme_bw)
```

It does not seem like any variables have a strong correlation aside from K and P. The next highest correlations have scores that are over .5 points lower. K and P have a correlation of 0.74, P and N have a correlation of -0.23, and temperature and humidity have a correlation of 0.21. This is not incredibly helpful for our modeling.